---
title: "stcp-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stcp-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6, fig.height=5,
  fig.align="center"
)
```

```{r setup}
library(stcp)
```


## Example
### 1. Sequential test

Suppose we have a stream of observations $X_1, X_2, \dots \in [0,1]$. Before an unknown change-point $\nu$, each pre-change observation has mean less than or equal to $0.2$. But after the change-point $\nu$, each post-change observation has mean larger than $0.2$.  

If $\nu$ is equal to either $0$ or $\infty$ then this setting is reduced to the following testing problem.

\[
H_0: \mu \leq 0.2~~\text{vs}~~H_1: \mu > 0.2
\]

For this example, we generate data stream from the beta distributions for simplicity.
However, the testing algorithm itself is nonparametric and does not rely on this specific choice of the distribution.

#### a) Under $H_0: \mu \leq 0.2$ 
```{r test_null_example}
# Set seed for the reproducibility.
set.seed(1) 
# Load stcp package
library(stcp)
# Generate a stream of observations under H0.
# For simplicity, we use beta distributions to generate samples.
max_sample <- 1000L
x_vec <- rbeta(max_sample, 2, 8) # H_0 dist: B(2, 8)
plot(1:max_sample, x_vec, pch=20, 
     xlab = "n", ylab = expression('X'['n']), main = "Simulated Data (H0)")
```

Below, we compute a mixture of e-values for testing the hypothesis.
```{r test_null_run}
alpha <- 1e-2 # Level 0.01 test 
m_pre <- 0.2 # Upper bound of mean of the null distribution.
delta_lower <- 0.01  # Guess on the minimum gap between the null and alternative means

# Build stcp model to test hypothesis on bounded random variables.
test_model <- build_stcp_bounded(alpha, m_pre, delta_lower, is_test = TRUE)

# Run the test
test_run <- run_stcp(x_vec, test_model)

# Printing summary
print(test_run)

# Plot 
if (is.infinite(test_run$stopped_ind)) {
  plot(test_run)
} else {
  plot(test_run, main = paste0("H0 is rejected at n = ", test_run$stopped_ind))
}

```

As the log of mixture of e-values does not cross the threshold ($\log(1/\alpha)\sim `r round(log(1/alpha), 2)`$), 
there is no enough evidence to reject the null hypothesis so far. 


#### b) Under $H_1: \mu_1 > 0.2$ 

Now, consider the case where the alternative hypothesis is true and the data generated from a distribution with $\mu > 0.2$.

```{r test_alter_example}
# Generate a stream of observations under H1
another_x_vec <- rbeta(max_sample, 2, 7) # H_1 dist: B(2,7)
plot(1:max_sample, another_x_vec, pch=20, 
     xlab = "n", ylab = expression('X'['n']), main = "Simulated Data (H1)")

```

Below, we use the same model to compute a mixture of e-values for testing the hypothesis.
```{r test_alter_run}
# Run the same stcp model used before but on the new data
test_run2 <- run_stcp(another_x_vec, test_model)

# Printing summary
print(test_run2)

# Plot 
if (is.infinite(test_run2$stopped_ind)) {
  plot(test_run2)
} else {
  plot(test_run2, main = paste0("H0 is rejected at n = ", test_run2$stopped_ind))
}
```

As the log of mixture of e-values crosses the threshold ($\log(1/\alpha)\sim `r round(log(1/alpha), 2)`$), 
we can reject the null at $n = `r test_run2$stopped_ind`$ at the earliest. 


#### c) Switching null and alternative.

Now, consider the following swapped hypotheses.
\[
H_0: \mu \geq 0.2~~\text{vs}~~H_1: \mu < 0.2
\]

To test the swapped hypothesis above, we can simply define a new random variable $Y := 1-X$ with $\theta := \mathbb{E}Y = 1-\mu$ and test the corresponding hypotheses.

\[
H_0: \theta \leq 0.8~~\text{vs}~~H_1: \theta > 0.8
\]

Based on the new random variable $Y$ and corresponding hypotheses, we can follow the same procedure above to conduct a valid sequential test.

```{r, opposite_direction}
theta_pre <- 0.8 # Upper bound of mean of the null distribution.
delta_lower <- 0.01  # Guess on the minimum gap between the null and alternative means

# Build stcp model to test hypothesis on bounded random variables.
test_model_op_direct <- build_stcp_bounded(alpha, theta_pre, delta_lower, is_test = TRUE)

# Run the test
# It is IMPORTANT to feed the new random variable (Y = 1-X)
test_run3 <- run_stcp(1-another_x_vec, test_model_op_direct) 

# Printing summary
print(test_run3)

# Plot 
if (is.infinite(test_run3$stopped_ind)) {
  plot(test_run3)
} else {
  plot(test_run3, main = paste0("H0 is rejected at n = ", test_run3$stopped_ind))
}

```
As the data was generated in the switched null hypothesis ($\mu \geq 0.2$), there is no enough evidence to reject the switched null hypothesis so far. 

Instead of defining a new random variable $Y$, the \code{build_stcp_bounded} function supports \code{is_flipped} argument to ``flip'' the original random variable $X$ into $Y := 1-X$ automatically.
```{r, opposite_direction2}
# Use is_flipped = TRUE to conduct the test for the swapped hypothesis.
test_model3b <- build_stcp_bounded(alpha, m_pre, delta_lower, is_test = TRUE, is_flipped = TRUE)


# Both models have same lambdas and omegas
all.equal(test_model_op_direct$lambda, test_model3b$lambda)
all.equal(test_model_op_direct$omega, test_model3b$omega)

# Run the test
test_run3b <- run_stcp(another_x_vec, test_model3b)

# Run results are also same
all.equal(test_run3$log_mix_e_vec, test_run3b$log_mix_e_vec)


# Printing summary
print(test_run3b)


# Plot 
if (is.infinite(test_run3b$stopped_ind)) {
  plot(test_run3b)
} else {
  plot(test_run3b, main = paste0("H0 is rejected at n = ", test_run3b$stopped_ind))
}


```


#### d) Two-sided test

Finally, let's examine how to conduct the following two-sided test.


\[
H_0: \mu = 0.2~~\text{vs}~~H_1: \mu \neq 0.2
\]

In this case, we need to combine two 'one-sided' tests into a single 'two-sided' test. In general, we can combine any two test runs into a single test run. The combined test run controls type-1 error under the intersection of two null hypotheses, which is equal to $\mu = 0.2$ in this example. If two tests share a common input then we can also combine two stcp models into a single model and then run a test based on the combined model. 

```{r, two_side_null}
# Combine two existing runs
test_run4 <- combine_stcp_run(test_run2, test_run3b)

# Or we can first combine models and re-run the model. 
# Note this method assumes both models take the same input. Thus we cannot use
# test_model_op_direct but must use test_model3b

combined_model <- combine_stcp(test_model, test_model3b)

test_run4b <- run_stcp(another_x_vec, combined_model)

# Two run results are same.
all.equal(test_run4$log_mix_e_vec, test_run4b$log_mix_e_vec)

# Printing summary
print(test_run4)

# Plot 
if (is.infinite(test_run4$stopped_ind)) {
  plot(test_run4)
} else {
  plot(test_run4, main = paste0("H0 is rejected at n = ", test_run4$stopped_ind))
}
plot(test_run2, lty = 2, col = 2, add = TRUE)
plot(test_run3, lty = 2, col = 3, add = TRUE)
```



### 2. Change-point detection

Now, back to the change-point detection setup. Recall that we have a stream of observations $X_1, X_2, \dots \in [0,1]$. Before an unknown change-point $\nu$, each pre-change observation has mean less than or equal to $0.2$. But after the change-point $\nu$, each post-change observation has mean larger than $0.2$. As same as the test example above, throughout change-point detection examples, we generate data stream from the beta distributions for simplicity. However, the testing algorithm itself is nonparametric and does not rely on this specific choice of the distribution. Below, we set the change-point $\nu = 500$ and generate 500 pre-change observations from $B(2,8)$. From time $\nu+1$, we also have 500 post-change observations from $B(2,6)$. 


```{r example}
# Generate a stream of observations
# For simplicity, we use beta distributions to generate samples
max_sample <- 1000L
nu <- 500L
a_pre <- 2
b_pre <- 8
a_post <- 2
b_post <- 6
x_vec <- numeric(max_sample)
x_vec[1:nu] <- rbeta(nu, a_pre, b_pre) # Pre-change dist
x_vec[seq(nu+1, max_sample)] <- rbeta((max_sample - nu), a_post, b_post) # Post-change dist
m_pre <- a_pre / (a_pre + b_pre)
m_post <- a_post / (a_post + b_post)
plot(1:max_sample, x_vec, pch=20, 
     xlab = "n", ylab = expression('X'['n']), main = "Simulated Data (v = 500)")
```

We compute a mixture e-detectors by which we can infer when the change of the mean happened.
```{r stcp}
alpha <- 1e-3 # Inverse of target ARL
m_pre <- m_pre # Upper bound of mean of pre-change observations
delta_lower <- 0.01  # Guess on the minimum gap between pre- and post-change means
stcp_model <- build_stcp_bounded(alpha, m_pre, delta_lower)

# Compute mixture of SR-type e-detectors.
mix_SR_bounded <- run_stcp(x_vec, stcp_model, is_SR_type = FALSE)
mix_SR_stop <- mix_SR_bounded$stopped_ind

# Printing summary
print(mix_SR_bounded)


# Plot 
plot(mix_SR_bounded, main = paste0("v = ", nu, "; v_hat = ",mix_SR_stop))
```


```{r example2}
plot(1:max_sample, x_vec, pch=20, 
     xlab = "n", ylab = expression('X'['n']), main = "Simulated Data with the detected CP")
lines(x = c(0,nu), y = c(m_pre,m_pre), col = 2, lwd = 2)
lines(x = c(nu,max_sample), y = c(m_post,m_post), col = 2, lwd =2 )
abline(v = nu, col = 1, lty = 2)
abline(v = mix_SR_stop, col = 2, lty = 2, lwd = 2)

```

<!--Use this commend to rebuild vignettes devtools::build_vignettes() -->
