---
title: "stcp-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stcp-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stcp)
```


## Example
### 1. Sequential test

Suppose we have a stream of observations $X_1, X_2, \dots \in [0,1]$. Before an unknown change-point $\nu$, each pre-change observation has mean less than or equal to $0.3$. But after the change-point $\nu$, each post-change observation has mean larger than $0.3$.  

If $\nu$ is equal to either $0$ or $\infty$ then this setting is reduced to the following testing problem.

\[
H_0: \mu \leq 0.2~~\text{vs}~~H_1: \nu > 0.2
\]

For this example, we generate data stream from Beta distributions for simplicity.
However, the testing algorithm itself is nonparametric and does not rely on this specific choice of the distribution.

#### a) Under $H_0$ 
```{r test_null_example}
set.seed(1)
library(stcp)
# Generate a stream of observations
# For simplicity, we use beta distributions to generate samples
max_sample <- 1000L
x_vec <- rbeta(max_sample, 2, 8) # H_0 dist: B(1, 3)
plot(1:max_sample, x_vec, pch=20, 
     xlab = "n", ylab = "X_n", main = "Simulated Data (H0)")
```

Below, we compute a mixture e-values for testing the hypothesis.
```{r test_null_run}
alpha <- 1e-2 # Level 0.01 test 
m_pre <- 0.2 # Upper bound of mean of the null distribution.
delta_lower <- 0.01  # Guess on the minimum gap between the null and alternative means
test_model <- build_stcp_bounded(alpha, m_pre, delta_lower, is_test = TRUE)

# Compute mixture of SR-type e-detectors.
mix_SR_test <- run_stcp(x_vec, test_model)
mix_SR_stop <- mix_SR_test$stopped_ind

# Printing summary
print(mix_SR_test)

# Plot 
plot(mix_SR_test)
```
As the log of mixture of e-values does not cross the threshold ($\log(1/\alpha)\sim `r round(log(1/alpha), 2)`$), 
there is no enough evidence to reject the null hypothesis so far. 


#### b) Under $H_1$ 

```{r test_alter_example}
# Generate a stream of observations
# For simplicity, we use beta distributions to generate samples
max_sample <- 1000L
x_vec <- rbeta(max_sample, 2, 7) # H_1 dist: B(2,3)
plot(1:max_sample, x_vec, pch=20, 
     xlab = "n", ylab = "X_n", main = "Simulated Data (H1)")

```

Below, we use the same model to compute a mixture e-values for testing the hypothesis.
```{r test_alter_run}
# Compute mixture of SR-type e-detectors.
mix_SR_test <- run_stcp(x_vec, test_model)
mix_SR_stop <- mix_SR_test$stopped_ind

# Printing summary
print(mix_SR_test)

# Plot 
plot(mix_SR_test)
```
As the log of mixture of e-values crosses the threshold ($\log(1/\alpha)\sim `r round(log(1/alpha), 2)`$), 
we can reject the null at $n = `r mix_SR_test$stopped_ind`$ at the earliest. 


### 2. Change-point detection


```{r example}
# Generate a stream of observations
# For simplicity, we use beta distributions to generate samples
max_sample <- 1000L
nu <- 500L
x_vec <- numeric(max_sample)
x_vec[1:nu] <- rbeta(nu, 2, 8) # Pre-change dist: B(2, 8)
x_vec[seq(nu+1, max_sample)] <- rbeta((max_sample - nu), 2, 7) # Post-change dist: B(2, 7)
m_pre <- 2 / (2+8)
m_post <- 2 / (2+7)
plot(1:max_sample, x_vec, pch=20, 
     xlab = "n", ylab = "X_n", main = "Simulated Data")
```

Below, we compute a mixture e-detectors by which we can infer when the change of the mean happened.
```{r stcp}
alpha <- 1e-3 # Inverse of target ARL
m_pre <- 0.2 # Upper bound of mean of pre-change observations
delta_lower <- 0.01  # Guess on the minimum gap between pre- and post-change means
stcp_model <- build_stcp_bounded(alpha, m_pre, delta_lower)

# Compute mixture of SR-type e-detectors.
mix_SR_bounded <- run_stcp(x_vec, stcp_model, is_SR_type = FALSE)
mix_SR_stop <- mix_SR_bounded$stopped_ind

# Printing summary
print(mix_SR_bounded)

# Plot 
plot(mix_SR_bounded, main = paste0("v = ", nu, "; v_hat = ",mix_SR_stop))
```


```{r example2}
plot(1:max_sample, x_vec, pch=20, 
     xlab = "n", ylab = "X_n", main = "Simulated Data with the detected CP")
lines(x = c(0,nu), y = c(0.3,0.3), col = 2, lwd = 2)
lines(x = c(nu,max_sample), y = c(0.4,0.4), col = 2, lwd =2 )
abline(v = nu, col = 1, lty = 2)
abline(v = mix_SR_stop, col = 2, lty = 2, lwd = 2)

```

<!--devtools::build_vignettes() -->
